---
title: "Average vs Median comparison"
output: rmarkdown::github_document
always_allow_html: yes
---

```{r echo=FALSE, include = FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

### Notebook objective
- Generate visualizations to show non-technical stakeholders how an underlying data distribution influences an average or median metric

### Metrics
- Median: find the  middle point (50th percentile) in a set of numbers
- Average: add a set of numbers up then divide by the length of the set

### Generate example data
- Simulated data for illustration purposes only
- Fake data for 3 different video games

```{r}
library(tidyverse)
library(kableExtra)

### generate example data
left_long_tail <- tibble(label="left_skew_shape",
                         nums = rbeta(10000,5,0.5)*1000) %>%
  mutate(avg = mean(nums),
         median = median(nums))

right_long_tail <- tibble(label="right_skew_shape",
                          nums = rbeta(10000,0.5,5)*1000) %>%
  mutate(avg = mean(nums),
         median = median(nums))

similar_tails  <- tibble(label="symmetric_shape",
                         nums = rbeta(10000,5,5)*1000) %>%
  mutate(avg = mean(nums),
         median = median(nums))

example_data <- bind_rows(list(left_long_tail, 
                               right_long_tail, 
                               similar_tails))
```

### Takeways
- Compared to the median, the average is pulled toward the tail of the underlying distribution
- When the underlying distribution has a symmetric shape then average and median will land close together
- When the average is considerably less than the median it hints at a left skew data shape
- When the average is considerably larger than the median it hints at a right skew data shape

```{r}
example_data %>%
  ggplot(aes(x=nums)) +
  geom_histogram(binwidth=20) +
  geom_vline(aes(xintercept=avg, color="average"), linetype = "dashed", size=1) +
  geom_vline(aes(xintercept=median, color="median"), size=1) +
  facet_wrap(. ~ label, ncol=1) +
  labs(title="Average vs median metric comparisons for different data shapes",
       y="Player Count",
       x="Points Scored") +
  scale_color_manual(name = "Metrics", values = c(median = "dodgerblue", average = "orange")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size=12),
        legend.position = "top",
        legend.title.align = 0.5)  + 
  guides(color = guide_legend(override.aes = list(size=5)))
```

```{r}
example_data %>%
      group_by(dataset_label = label) %>%
      summarise(average_points_scored = round(mean(nums),0),
                median_points_scored = round(median(nums),0)) %>%
      kbl() %>%
      kable_classic(full_width = F, html_font = "Cambria")
```


